Conditional Generation of Variation and Mean Fitness Increase: \\ Necessity is the Mother of Genetic Invention

### 2017-11-26 ms.tex
revised mutational loss section after MWF comments

Note that if the modifier has the same effect on all types,
then we can substitute $C = \mu I$ (with $\mu > 0$) in Eq.~\ref{eq:model_equilibrium}, and proceeding as in Eq.~\ref{eq:mfp_K}, we have
\begin{equation}\label{eq:mutational_loss}
\frac{\partial \hat{\bar{w}}}{\partial \mu} = 
-\frac{1}{\mu}\Big(\sum_{k=1}^{n} \hat u_k \hat v_k w_k - \hat{\bar{w}} \Big) = -\frac{1}{\mu} G.
\end{equation}
This relationship has been derived by~\citet[eq.~24]{Hermisson2002}.
They called $G$ the \emph{mutational loss}: it is the difference between the \emph{ancestral mean fitness} $(\sum_{k=1}^{n} \hat u_k \hat v_k w_k)$ and the \emph{stable population mean fitness} $(\hat{\bar{w}})$, .

If an invading modifier allele $M$ reduces the transition rate compared to the resident allele $m$, i.e. $C_k = \hat \mu$ and $\tilde{C}_k = \mu$ for all $k$ in Eq.~\ref{eq:Lex} and $\mu < \hat \mu$, then $\cl_{ex}$ becomes
\begin{equation}
\cl_{ex} = ((1 - \mu)I + \mu M) D /\hat{\bar w},
\end{equation} 
and we can we can expect that $M$ will be favored by natural selection.
That is, the \emph{unified reduction principle} is in effect~\cite[eqs. 65, 72]{Altenberg2017}.
Thus, we can conclude that the \emph{mutational loss} $G$ is positive and that the \emph{ancestral mean fitness} is higher then the \emph{stable population mean fitness}.

>>>>

Note that if the modifier has the same effect on all types,
then we can substitute $C = \mu I$ (with $\mu > 0$) in Eq.~\ref{eq:model_equilibrium}, and proceeding as in Eq.~\ref{eq:mfp_K}, we find a relationship previously described by~\citet[eq.~24]{Hermisson2002},
\begin{equation}\label{eq:mutational_loss}
\frac{\partial \hat{\bar{w}}}{\partial \mu} = 
-\frac{1}{\mu}\Big(\sum_{k=1}^{n} \hat u_k \hat v_k w_k - \hat{\bar{w}} \Big) = -\frac{1}{\mu} G,
\end{equation}
where $G$, equal to the difference between the \emph{ancestral mean fitness} $(\sum_{k=1}^{n} \hat u_k \hat v_k w_k)$ and the \emph{stable population mean fitness} $(\hat{\bar{w}})$, is called the \emph{mutational loss}~\citep{Hermisson2002}.

If an invading modifier allele $M$ reduces the transition probability compared to the resident allele $m$, i.e. $C_k = \hat \mu$ and $\tilde{C}_k = \mu$ for all $k$ in Eq.~\ref{eq:Lex} and $\mu < \hat \mu$, then $\cl_{ex}$ becomes
\begin{equation}
\cl_{ex} = ((1 - \mu)I + \mu M) D /\hat{\bar w},
\end{equation} 
and we can we can expect that $M$ will be favored by natural selection.
That is, the \emph{unified reduction principle} is in effect~\cite[eqs. 65, 72]{Altenberg2017}.
Thus, we can conclude that the \emph{mutational loss} $G$ is positive and that the \emph{ancestral mean fitness} is higher then the \emph{stable population mean fitness}.
